---
title: "p8105_hw2_qc2337"
author: "cathy"
date: "2024-10-02"
output: github_document
---
Load all the package needed

```{r setup, echo = FALSE, message=FALSE}
library(tidyverse)
```

## Problem 1

 Read and clean data

```{r}
 nyc_transit_df =
   read_csv("NYC_Transit_Subway_Entrance_And_Exit_Data.csv")|> 
   janitor::clean_names() |> 
   select(line, station_name, station_latitude, station_longitude, 
          route1:route11, entrance_type, entry, vending, ada) |> 
   mutate(
      entry = case_match(
       entry,
        "YES" ~ TRUE,
        "NO" ~ FALSE
    )
  )
```

### Short description for the dataset
In this 'nyc_transit_df' dataset, there are 19 variables including line, station_name, station_latitude, station_longitude, route1 to route 11, entrance_type, entry, vending, and ada.

Steps for data cleaning include: 
 1. Make all variables lowcase.
 2. Select the variables I need.
 3. Using 'mutate' to convert the entry variable from a character to a logical variable.

I got the dataset with 1,868 × 19. These data are not tidy because there are 11 columns to contain 'route', it should be included in 1 column.

### Answer the following questions using these data:

1. How many distinct stations are there? 

```{r}
 nycstation_df = 
   nyc_transit_df |> 
   distinct(line, station_name)

 nycstation_df
```

Create a new variable 'nycstation_df', using 'distinct' we got 465 distinct stations.

2. How many stations are ADA compliant?

```{r}
 ada_station_df =
   nyc_transit_df |> 
     select(ada, line, station_name) |>
     filter(ada) |> 
     distinct(line, station_name)

 ada_station_df 
```

 84 stations are ADA compliant.

3. What proportion of station entrances / exits without vending allow entrance?

```{r}
 no_vending_entry = 
   nyc_transit_df |> 
      filter(vending == "NO", entry == TRUE)

 no_vending_station =
   nyc_transit_df |> 
   filter(vending == "NO")
 
 
proportion = nrow(no_vending_entry) / nrow(no_vending_station)

proportion

```

The proportion of station entrances / exits without vending allow entrance is 0.3770492. 


### Reformat data

Make route number and route name distinct variables
and answer 2 questions

```{r}
 nyc_routes =
   nyc_transit_df |> 
   mutate(across(route1:route11, as.character)) |> 
   pivot_longer(
    cols = route1:route11,
    names_to = "route_number",
    values_to = "route_value",
    names_prefix = "route"
) |> 
   filter(route_value == "A", ada == TRUE) |> 
   distinct(line, station_name)
  
 nyc_routes
```

60 distinct stations serve the A train.
Of the stations that serve the A train, 17 are ADA compliant.


## Problem 2




