---
title: "p8105_hw2_qc2337"
author: "cathy"
date: "2024-10-02"
output: github_document
---
Load all the package needed

```{r setup, echo = FALSE, message=FALSE}
library(tidyverse)
```

## Problem 1

Read and clean data

```{r}
 nyc_transit_df =
   read_csv("NYC_Transit_Subway_Entrance_And_Exit_Data.csv")|> 
   janitor::clean_names() |> 
   select(line, station_name, station_latitude, station_longitude, 
          route1:route11, entrance_type, entry, vending, ada) |> 
   mutate(
      entry = case_match(
       entry,
        "Yes" ~ TRUE,
        "No" ~ FALSE
    )
  )
```

In this 'nyc_transit_df' dataset, there are 19 variables including line, station_name, station_latitude, station_longitude, route1 to route 11, entrance_type, entry, vending, and ada.

Steps for data cleaning include: 
 1. Make all variables lowcase.
 2. Select the variables I need.
 3. Using 'mutate' to convert the entry variable from a character to a logical variable.

I got the dataset with 1,868 × 19. These data are not tidy because there are 11 columns to contain 'route', it should be included in 1 column.
